---
title: "Final Project"
output: html_notebook
---
# load library
```{r}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(vip)
library(parallel)
library(doParallel)
library(embed)
library(rpart.plot) 
library(rpart) 
library(kableExtra)
library(GGally)
library(vip) 
library(fastshap)  
library(MASS)
library(caret)
library(glmnet)
library(reshape2)

```

# import data
```{r}
car_loan <- read_csv("car_loan.csv") %>%
  clean_names() 
head(car_loan)
```

# skim 
```{r}
car_loan %>%
  skim()
```

# Mutate variables
## transform average_acct_age and credit_history_length
```{r}
car_loan %>%
  mutate(acct_yr=  chartr('yrs','   ',substr(average_acct_age,0,2))) %>%
  mutate(acct_yr= as.numeric(str_replace_all(acct_yr, " ", "")))%>%
  mutate(acct_mth=  chartr('mon','   ',substr(average_acct_age,6,8))) %>%
  mutate(acct_mth= as.numeric(str_replace_all(acct_mth, " ", "")))%>%
  mutate(average_acct_month = acct_yr * 12 + acct_mth) %>%
  mutate(credit_yr=  chartr('yrs','   ',substr(credit_history_length,0,2))) %>%
  mutate(credit_yr= as.numeric(str_replace_all(credit_yr, " ", "")))%>%
  mutate(credit_mth=  chartr('mon','   ',substr(credit_history_length,6,8))) %>%
  mutate(credit_mth= as.numeric(str_replace_all(credit_mth, " ", "")))%>%
  mutate(credit_history_month = credit_yr * 12 + credit_mth) %>%
  dplyr::select(-acct_yr,-acct_mth,-average_acct_age,-credit_yr,-credit_mth,-credit_history_length) ->car_loan

```

## date into days
```{r}
car_loan %>%
  mutate(date_of_birth=chartr('/','-',date_of_birth)) %>%
  mutate(date_of_birth=as.Date(date_of_birth,format='%d-%m-%Y')) %>%
  mutate(birth_days=as.numeric(as.Date("2021-01-01")-date_of_birth))%>%
  dplyr::select(-date_of_birth)->car_loan
```

## as factor
```{r}
car_loan %>%
  mutate(loan_default=as.factor(loan_default))->car_loan
```


# Data Exploration
## Explore target
```{r}
car_loan %>%
  count(loan_default) %>%
  mutate(pct = n/sum(n)) -> default_rate

default_rate %>%
  ggplot(aes(x=loan_default, y=pct)) +
  geom_col() +
  geom_text(aes(label=pct) ,color="red") + 
  labs(title="Default Rate")
```

## Explore numerics
```{r}
boxplot<- function(a,b){
  ggplot(car_loan, aes(x={{a}},y=loan_default,group=loan_default)) + geom_boxplot() +labs(title = b)
}

boxplot(disbursed_amount,"disbursed_amount")
boxplot(asset_cost,"asset_cost")
boxplot(ltv,"ltv")
boxplot(branch_id,"branch_id")
boxplot(supplier_id,"supplier_id")
boxplot(manufacturer_id,"manufacturer_id")
boxplot(current_pincode_id,"current_pincode_id")
boxplot(state_id,"state_id")
boxplot(employee_code_id,"employee_code_id")
boxplot(mobile_no_avl_flag,"mobile_no_avl_flag")
boxplot(aadhar_flag,"aadhar_flag")
boxplot(pan_flag,"pan_flag")
boxplot(voter_id_flag,"voter_id_flag")
boxplot(driving_flag,"driving_flag")
boxplot(passport_flag,"passport_flag")
boxplot(perform_cns_score,"perform_cns_score")
boxplot(pri_no_of_accts,"pri_no_of_accts")
boxplot(pri_active_accts,"pri_active_accts")
boxplot(pri_overdue_accts,"pri_overdue_accts")
boxplot(pri_current_balance,"pri_current_balance")
boxplot(pri_sanctioned_amount,"pri_sanctioned_amount")
boxplot(pri_disbursed_amount,"pri_disbursed_amount")
boxplot(sec_no_of_accts,"sec_no_of_accts")
boxplot(sec_active_accts,"sec_active_accts")
boxplot(sec_overdue_accts,"sec_overdue_accts")
boxplot(sec_current_balance,"sec_current_balance")
boxplot(sec_sanctioned_amount,"sec_sanctioned_amount")
boxplot(sec_disbursed_amount,"sec_disbursed_amount")
boxplot(primary_instal_amt,"primary_instal_amt")
boxplot(sec_instal_amt,"sec_instal_amt")
boxplot(new_accts_in_last_six_months,"new_accts_in_last_six_months")
boxplot(delinquent_accts_in_last_six_months,"delinquent_accts_in_last_six_months")
boxplot(no_of_inquiries,"no_of_inquiries")
boxplot(average_acct_month,"average_acct_month")
boxplot(credit_history_month,"credit_history_month")
boxplot(birth_days,"age")
```

## Explore character
```{r}
bar_plot <- function(a,colname){
    car_loan %>%
    ggplot(., aes(a)) + 
    geom_bar(aes(fill = loan_default), position = "fill") +
    labs(title = paste("Barplot for",colname), x = colname, y="percent")+
    geom_hline(yintercept = 0.21707)+
    theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5))
}
bar_plot(car_loan$employment_type,"employment_type")
bar_plot(car_loan$perform_cns_score_description,"perform_cns_score_description")
```

## Explore correlation
```{r}
library(corrplot)
cormat <- cor(select_if(car_loan%>%na.omit(),is.numeric))
round(cormat, 2)
corrplot(cormat,tl.cex  = 0.5)
```

# define train/test split 
```{r}
set.seed(42)
train_test_spit<- initial_split(car_loan, prop = 0.7)

train <- training(train_test_spit)
test  <- testing(train_test_spit)

train_cv_folds <- vfold_cv(train, v=5)
```

# define recipe
```{r}
recipe <- recipe(loan_default ~ ., 
                      data = train) %>%
  step_rm(unique_id,mobile_no_avl_flag) %>%
  step_novel(all_nominal_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())
```

# logistic regression full
```{r}
logistic_spec <- logistic_reg() %>%
  set_mode("classification") %>%
  set_engine("glm")

logistic_full_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(logistic_spec) %>%
  fit(train)

logistic_full_wf %>%
  pull_workflow_fit() %>%
  tidy() %>%
  mutate(across(is.numeric,round,3))

logistic_full_wf %>%
  pull_workflow_fit() %>%
  vip()

options(yardstick.event_first = FALSE)
predict(logistic_full_wf, train, type="prob") %>%
  bind_cols(predict(logistic_full_wf, train, type="class")) %>%
  bind_cols(train) ->logistic_train

logistic_train %>%
  metrics(loan_default, estimate = .pred_class, .pred_1)

predict(logistic_full_wf, test, type="prob") %>%
  bind_cols(predict(logistic_full_wf, test, type="class")) %>%
  bind_cols(test) -> logistic_test 

logistic_test %>%
  metrics(loan_default, estimate = .pred_class, .pred_1)
```

# reduced logistic
```{r}
recipe1 <- recipe(loan_default ~ ., 
                      data = train) %>%
  step_rm(unique_id,mobile_no_avl_flag,voter_id_flag,pri_sanctioned_amount,
          pri_disbursed_amount,sec_current_balance,primary_instal_amt,sec_instal_amt,
          new_accts_in_last_six_months) %>%
  step_novel(all_nominal_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())

logistic_spec <- logistic_reg() %>%
  set_mode("classification") %>%
  set_engine("glm")

logistic_full_wf1 <- workflow() %>%
  add_recipe(recipe1) %>%
  add_model(logistic_spec) %>%
  fit(train)

logistic_full_wf1 %>%
  pull_workflow_fit() %>%
  tidy() %>%
  mutate(across(is.numeric,round,3))

logistic_full_wf1 %>%
  pull_workflow_fit() %>%
  vip()

options(yardstick.event_first = FALSE)
predict(logistic_full_wf1, train, type="prob") %>%
  bind_cols(predict(logistic_full_wf1, train, type="class")) %>%
  bind_cols(train) ->logistic_train1

logistic_train1 %>%
  metrics(loan_default, estimate = .pred_class, .pred_1)

predict(logistic_full_wf1, test, type="prob") %>%
  bind_cols(predict(logistic_full_wf1, test, type="class")) %>%
  bind_cols(test) -> logistic_test1

logistic_test1 %>%
  metrics(loan_default, estimate = .pred_class, .pred_1)
```

# default decision tree
```{r}
tree_spec <- decision_tree(c=0) %>%
  set_mode("classification") %>%
  set_engine("rpart")

tree_wf1 <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(tree_spec) %>%
  fit(train)

tree_wf1 %>%
  pull_workflow_fit() %>%
  vip()

# -- deal w. the first event issue -- # 
options(yardstick.event_first = FALSE)
predict(tree_wf1, train, type="prob") %>%
  bind_cols(predict(tree_wf1, train, type="class")) %>%
  bind_cols(train) ->tree_train1

tree_train1%>%
  metrics(loan_default, estimate = .pred_class, .pred_1)

predict(tree_wf1, test, type="prob") %>%
  bind_cols(predict(tree_wf1, test, type="class")) %>%
  bind_cols(test) -> tree_test1

tree_test1 %>%
  metrics(loan_default, estimate = .pred_class, .pred_1)
```

## calculate cp
```{r}
set.seed(123)
tree1 <- train(loan_default ~ .,
                   data = bake_train, 
                  method="rpart",
                  trControl = trainControl("cv", number = 40),
                  tuneLength = 20)
plot(tree1)
tree1$bestTune
```

# Best cp tree
```{r}
tree_spec2 <- decision_tree(c=0.0001979358) %>%
  set_mode("classification") %>%
  set_engine("rpart")

tree_wf2 <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(tree_spec2) %>%
  fit(train)

tree_wf2 %>%
  pull_workflow_fit() %>%
  vip()

# -- deal w. the first event issue -- # 
options(yardstick.event_first = FALSE)
predict(tree_wf2, train, type="prob") %>%
  bind_cols(predict(tree_wf2, train, type="class")) %>%
  bind_cols(train) ->tree_train2
tree_train2%>%
  metrics(loan_default, estimate = .pred_class, .pred_1)

predict(tree_wf2, test, type="prob") %>%
  bind_cols(predict(tree_wf2, test, type="class")) %>%
  bind_cols(test) -> tree_test2
tree_test2 %>%
  metrics(loan_default, estimate = .pred_class, .pred_1)
```

#Random forest
```{r,message=FALSE,warning=FALSE}
cores <- parallel::detectCores()
print(cores)


forest_spec <- rand_forest(
  mtry=7,
  trees = 200,
  min_n = 10
  ) %>%
  set_mode("classification") %>%
  set_engine("ranger",
             num.threads = 8,  
             importance = "impurity")

forest_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(forest_spec) %>%
  fit(train)

forest_wf %>% 
  pull_workflow_fit() %>%
  vip(20)

```

## Evaluation 
```{r,message=FALSE,warning=FALSE}
options(yardstick.event_first = FALSE)
train_score_rf <-predict(forest_wf , train, type="prob") %>%
bind_cols(predict(forest_wf , train, type="class")) %>%
bind_cols(train)

test_score_rf <-predict(forest_wf , test, type="prob") %>%
bind_cols(predict(forest_wf , test, type="class")) %>%
bind_cols(test)
  
# -- metrics -- 
train_score_rf %>% 
  metrics(loan_default, estimate = .pred_class, .pred_1)
test_score_rf %>%
  metrics(loan_default, estimate = .pred_class, .pred_1)

# -- roc curve plot -- 
test_score_rf %>%
  mutate(model = "forest test") %>%
  bind_rows(train_score_rf %>%
              mutate(model="forest  train")) %>%
  group_by(model) %>%
  roc_curve(loan_default, .pred_1) %>%
  autoplot()
```

```{r,message=FALSE,warning=FALSE}

# --model spec w. tuning 
rf_model <- rand_forest(
    trees  = tune(),
    min_n = tune(),
   ) %>% 
      set_engine("ranger", importance = "impurity") %>% 
      set_mode("classification")

# -- workflow 
rf_workflow <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(rf_model) 



# --tuning grid -- random force 
tune_grid <- grid_random(trees(c(100,500)),
                         min_n(),
                          size = 5)
print(tune_grid)

# --parallel process 
all_cores <- detectCores(logical = TRUE)
sprintf("# of Logical Cores: %d", all_cores)
cl <- makeCluster(all_cores)
registerDoParallel(cl)

# -- train K times for each parameter -- 
rf_tuning_results <- rf_workflow %>% 
  tune_grid(
    resamples = train_cv_folds,
    grid = tune_grid,
    control = control_resamples(save_pred = TRUE)
    )

rf_tuning_results

```
## Review Tuning Results 
```{r}
## -- results of tuning -- 
rf_tuning_results %>% 
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>% 
  pivot_wider(names_from = .metric, values_from=c(mean, std_err))
```


## Visualize impact 

```{r}
## - visualize 
rf_tuning_results %>%
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>%
  ggplot(aes(trees, mean, )) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 

rf_tuning_results %>%
  collect_metrics()  %>%
  mutate_if(is.numeric, round,3) %>%
  ggplot(aes(min_n, mean, )) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 

```

## refitting workflow with "best" parameters

```{r}
rf_tuning_results %>%
  show_best("accuracy") %>%
  print()

rf_best <- rf_tuning_results %>%
  select_best("accuracy") 

print(rf_best)
rf_final_wf <- rf_workflow %>% 
  finalize_workflow(rf_best )

print(rf_final_wf)

rf_final_fit  <- rf_final_wf %>%
  fit(data = train) 
```

#variable of importance
```{r}
rf_final_fit %>% 
  extract_fit_parsnip() %>% 
  vip(20)
```

#predict
```{r}

scored_train_rft <-predict(rf_final_fit  , train, type="prob") %>%
bind_cols(predict(rf_final_fit  , train, type="class")) %>%
bind_cols(train)

 scored_test_rft <-predict(rf_final_fit  , test, type="prob") %>%
bind_cols(predict(rf_final_fit , test, type="class")) %>%
bind_cols(test)

  
```
#evaluation
```{r,message=FALSE,warning=FALSE}
calc_metrics <- function(data_set){
  data_set %>%
  conf_mat(loan_default, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="confusion matrix default") -> p 
  print(p)

data_set %>%
  accuracy(loan_default, estimate = .pred_class) %>%
  bind_rows(data_set %>%
  precision(loan_default, estimate = .pred_class)   ) %>%
  bind_rows(data_set %>%
  recall(loan_default, estimate = .pred_class)  )
}
```

#Evaluation result
```{r}
calc_metrics(scored_train_rft )
calc_metrics(scored_test_rft)
# -- roc curve plot -- 
scored_test_rft%>%
  mutate(model = "rf test") %>%
  bind_rows(scored_train_rft %>%
              mutate(model="rf train")) %>%
  group_by(model) %>%
  roc_curve(loan_default,.pred_1) %>%
  autoplot()

```


