---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(kableExtra)
library(GGally)
library(kableExtra) # -- make nice looking resutls when we knitt 
library(vip)        # --  tidymodels variable importance
library(fastshap)   # -- shapley values for variable importance 
library(rpart.plot) # -- plotting decision trees 
library(MASS)
library(caret)
library(rpart)
```

# import data
```{r}
org<- read_csv("organics.csv") %>% clean_names()
head(org)
neworg <- read_csv("new_organics.csv") %>% clean_names()
head(neworg)
```
# remove columns
```{r}
subset(org,select=c(dem_affl,dem_age,dem_cluster_group,
                    dem_gender,dem_reg,dem_tv_reg,
                    prom_class,prom_spend,prom_time,target_buy))->org
org
colnames(org)<-make.unique(names(org))
```

# skim organics
```{r}
org %>%skim()
neworg%>%skim()
```

# Explore variables
## Explore target
```{r}
org %>%
  ggplot(aes(x=target_buy)) +
  geom_histogram(stat="count") +
  labs(title = "Buy vs. Not Buy ")

org %>%
  group_by(target_buy) %>%
  summarize(n=n()) %>%
  ungroup() %>%
  mutate(pct = n/sum(n))

```

## Explore numerics
```{r}
org<-org%>%
  mutate(target_buy=as.factor(target_buy))
# -- comparative histogram
ggplot(org,aes(x=dem_affl))+
  geom_histogram(data=subset(org,target_buy==0),fill="red", alpha = 0.5, bins = 50) +
  geom_histogram(data=subset(org,target_buy==1),fill = "blue", alpha = 0.5,  bins = 50) +
  labs(title="Affluence Grade by target_buy = 0 (red) vs by target_buy = 1 (blue)")

ggplot(org,aes(x=dem_age))+
  geom_histogram(data=subset(org,target_buy==0),fill="red", alpha = 0.5, bins = 50) +
  geom_histogram(data=subset(org,target_buy==1),fill = "blue", alpha = 0.5,  bins = 50) +
  labs(title="Age by target_buy = 0 (red) vs by target_buy = 1 (blue)")

ggplot(org,aes(x=prom_spend))+
  geom_histogram(data=subset(org,target_buy==0),fill="red", alpha = 0.5, bins = 50) +
  geom_histogram(data=subset(org,target_buy==1),fill = "blue", alpha = 0.5,  bins = 50) +
  labs(title="Total amount spent in the store this year by target_buy = 0 (red) vs by target_buy = 1 (blue)")

ggplot(org,aes(x=prom_time))+
  geom_histogram(data=subset(org,target_buy==0),fill="red", alpha = 0.5, bins = 50) +
  geom_histogram(data=subset(org,target_buy==1),fill = "blue", alpha = 0.5,  bins = 50) +
  labs(title="Time as loyalty card member by target_buy = 0 (red) vs by target_buy = 1 (blue)")

# -- comparative boxplots
ggplot(org, aes(x=dem_affl, y=target_buy)) + geom_boxplot() +labs(title = "Affluence Grade")
ggplot(org, aes(x=dem_age, y=target_buy)) + geom_boxplot() +labs(title = "Age")
ggplot(org, aes(x=prom_spend, y=target_buy)) + geom_boxplot() +labs(title = "Total amount spent in the store this year")
ggplot(org, aes(x=prom_time, y=target_buy)) + geom_boxplot() +labs(title = "Time as loyalty card member")

```

## Explore characters
```{r}

char_explore <- function(col){
  org %>%
    ggplot(., aes(!!as.name(col))) + 
    geom_bar(aes(fill = target_buy), position = "fill") 
}

# -- for each character column if it doesnt equal customer id then create a chart
for (column in names(org %>% select_if (is_character))){
  chrt <- char_explore(column)
    print(chrt)
  
}
```
# Generate Models
## 0. Make Factors! 
```{r}
org %>%
    mutate_if(is.character, factor)-> org_prep
org_prep$target_buy <- as.factor(org_prep$target_buy)
head(org_prep)
```

## 1. Partition your data 70/30 (train / test split) 


```{r}
# -- set a random seed for repeatablity 
set.seed(43)

# -- performs our train / test split 
org_split <- initial_split(org_prep, prop = 0.7)

# -- extract the training data 
org_train <- training(org_split)
# -- extract the test data 
org_test <- testing(org_split)

sprintf("Train PCT : %1.2f%%", nrow(org_train)/ nrow(org_prep) * 100)
sprintf("Test  PCT : %1.2f%%", nrow(org_test)/ nrow(org_prep) * 100)

```

## 2. Recipe

```{r}
org_recipe <- recipe(target_buy ~ ., data = org_train) %>%
  step_mutate(dem_gender=replace_na(dem_gender,"U")) %>%
  step_modeimpute(all_nominal(), -all_outcomes()) %>%
  step_medianimpute(all_numeric()) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  prep()
org_recipe
```


## 3. Bake 

```{r}
bake_train <- bake(org_recipe, new_data = org_train)
bake_test  <- bake(org_recipe, new_data = org_test)
```


## 4.Fit
### 4.1 Default Decision tree

```{r}
org_tree <- decision_tree(mode="classification") %>%
                  set_engine("rpart") %>%
                  fit(target_buy ~ ., data = bake_train)

org_tree

rpart.plot(org_tree$fit, roundint=FALSE, extra=3)

```

### Decision tree 2
```{r}
set.seed(123)
org_tree1 <- train(target_buy ~ .,
                   data = bake_train, 
                  method="rpart",
                  trControl = trainControl("cv", number = 40),
                  tuneLength = 20)
plot(org_tree1)
org_tree1$bestTune
```

```{r}
org_tree2 <- decision_tree(mode="classification",
                            cost_complexity = 0.0011,
                            tree_depth = 6,
                            min_n = 2) %>%
                  set_engine("rpart") %>%
                  fit(target_buy ~ ., data=bake_train)

org_tree2$fit

options(scipen = 0)

rpart.plot(org_tree2$fit, roundint=FALSE, extra=3)

```

## 4.2 Logistic Regression
### 4.2.1 full model
```{r}
logistic_glm <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(target_buy ~ ., data = bake_train)

tidy(logistic_glm) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)

```

### 4.2.2 Stepwise Model
```{r}
steplog <- glm(target_buy ~ ., data = bake_train, family=binomial(link="logit"))
step <- stepAIC(steplog, direction="both")
summary(step)

```



```{r}
## -- Use tidymodel framework to fit and evaulate reduced model
org_steprecipe <- recipe(target_buy ~ dem_affl + dem_age + dem_gender , data = org_train) %>%
  step_medianimpute(all_numeric()) %>%
  prep()
org_steprecipe

# -- apply new recipe 
bake_steptrain <- bake(org_steprecipe, new_data = org_train)
bake_steptest  <- bake(org_steprecipe, new_data = org_test)

logistic_step1 <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(target_buy ~ ., data = bake_steptrain)

## -- check out your parameter estimates ... 
tidy(logistic_step1) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)
```

## 5. Prep for Evaluation 
### 5.1 Decision tree 1
```{r}

# -- training 
predict(org_tree, bake_train, type = "prob") %>%
  bind_cols(.,predict(org_tree, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_tree

head(scored_train_tree)

# -- testing 
predict(org_tree, bake_test, type = "prob") %>%
  bind_cols(.,predict(org_tree, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_tree

head(scored_test_tree)
```

### 5.2 Decision tree 2
```{r}
# -- training 
predict(org_tree2, bake_train, type = "prob") %>%
  bind_cols(.,predict(org_tree2, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_tree2

head(scored_train_tree2)

# -- testing 
predict(org_tree2, bake_test, type = "prob") %>%
  bind_cols(.,predict(org_tree2, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_tree2

head(scored_test_tree2)
```


### 5.3 logistic regression full model
```{r}
# -- training 
predict(logistic_glm, bake_train, type = "prob") %>%
  bind_cols(.,predict(logistic_glm, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_glm

head(scored_train_glm)

# -- testing 
predict(logistic_glm, bake_test, type = "prob") %>%
  bind_cols(.,predict(logistic_glm, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_glm

head(scored_test_glm)
```

### 5.4 logistic regression reduced model
```{r}
# -- training predictions from reduced model
predict(logistic_step1, bake_steptrain, type = "prob") %>%
  bind_cols(.,predict(logistic_step1, bake_steptrain)) %>%
  bind_cols(.,bake_steptrain) -> scored_train_step1

head(scored_train_step1)

# -- testing predictions from reduced model
predict(logistic_step1, bake_steptest, type = "prob") %>%
  bind_cols(.,predict(logistic_step1, bake_steptest)) %>%
  bind_cols(.,bake_steptest) -> scored_test_step1

head(scored_test_step1)
```


## 6. Evaluate

### 6.1 Decision tree 1
```{r}
options(yardstick.event_first = FALSE)
# -- AUC: Train and Test 
scored_train_tree %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_tree %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 

# -- Variable Importance top 10 features  
org_tree %>%
  vip(num_features = 5)

# -- ROC Charts 
scored_train_tree %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_tree %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()


# -- Confustion Matricies  
scored_train_tree %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_tree %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")

```

### 6.2 decision tree 2
```{r}
options(yardstick.event_first = FALSE)
# -- AUC: Train and Test 
scored_train_tree2 %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows(scored_test_tree2 %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 

# -- Variable Importance top 10 features  
org_tree %>%
  vip(num_features = 5)

# -- ROC Charts 
scored_train_tree2 %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_tree2 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()


# -- Confustion Matricies  
scored_train_tree2 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_tree2 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")

```

### 6.3 logistic regression full model
```{r}
# -- AUC: Train and Test 
scored_train_glm %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_glm %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") ) 

# -- Variable Importance top 10 features  
logistic_glm %>%
  vip(num_features = 10)

# -- ROC Charts 
scored_train_glm %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_glm %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()

# -- Confustion Matricies  
scored_train_glm %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_glm %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")
```

### 6.4 reduced model
```{r}
# -- Evaluate Reduced Model
# -- AUC: Train and Test 
scored_train_step1 %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_step1 %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  )

options(yardstick.event_first = FALSE)
# -- ROC Charts 
scored_train_step1 %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_step1 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()

# -- Confustion Matricies  
scored_train_step1 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_step1 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")
```

# Fit in new chart
## 1. subset chart
```{r}
subset(neworg,select=c(dem_affl,dem_age,dem_cluster_group,
                       dem_gender,dem_reg,dem_tv_reg,prom_class,prom_spend,prom_time))->neworg
colnames(neworg)<-make.unique(names(neworg))
neworg
  
```
## 2. Bake
```{r}
bake_neworg  <- bake(org_recipe, new_data = neworg) %>%
  mutate(dem_cluster_group_U=0,dem_tv_reg_C.Scotland=0)
```

## 3. predict
```{r}
predict(org_tree2, bake_neworg, type = "prob") %>%
  bind_cols(.,predict(org_tree2, bake_neworg)) %>%
  bind_cols(.,bake_neworg) -> scored_neworg_tree

head(scored_neworg_tree)
```
## 4.output table
```{r}
write.table(scored_neworg_tree,"new_organics_prediction.csv",row.names=FALSE,col.names=TRUE,sep=",")
```

